How Household Income Impacts Political Views, Focusing on Mr. Donald Trump, Yielded Inconclusive Results

### 2024-02-15 outputs/paper/paper.qmd
Final Paper updates

---
title: "How Household Income Impacts Political Views, Focusing on Mr. Donald Trump, Yielded Inconclusive Results"
author: 
  - Gavin Crooks
  - Samarth Rajani
thanks: "Code and data are available at: LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media"
date: today
date-format: long
abstract: "This study explores the relationship between household income, ethnicity, and political preferences using data from 'The Welfare Effects of Social Media.' Despite inconclusive findings challenging the presumed impact of income on political views, our innovative metric, measuring respondents' propensity to follow Donald Trump, did not provide conclusive results. This introduces new avenues in political science and statistics, encouraging further exploration into understanding the dynamics of these relationships. The study contributes to the ongoing discourse on the complex relationships between income, ethnicity, and political inclinations." 
format: pdf
number-sections: true
toc: true
bibliography: references.bib

@@

Our data has been obtained from ‘The Welfare Effects of Social Media’ [@paper] .
Our code is supported by the following packages [@citeR] [@tidy] [@her] [@knitr]

@@

The data for this paper was collected from the replication package of the paper ‘The Welfare Effects of Social Media’[@paper] The authors from that data had collected this data themselves using an online survey platform called Qulatrics, inquiring about personal information such as personal names, IP addresses, extent of by which the subjects follow politics, etc. Notably the dataset would contain a lot of confidential information, that if released in the replication package would cause ethical problems. As a result, the authors included the de-identified versions of their data collected, which was used in our analysis too.

@@

Another variable of use is the extent to which the subjects follow Mr. Donald Trump, leader of the Republican Party. The possible responses were ‘Not at all closely', 'Somewhat closely,’ ‘Rather closely’ and ‘Very Closely,’ and the respondents could select one of these options which will become our measure of measuring subscription to Republican ideas. This data is represented using a pie chart in @fig-figure2. This variable faces many problems as this categorical scale is not consistent across respondents. To be specific we mean someone who identifies as someone who does follows Not at all closely can be following Mr. Donald Trump more than someone who identifies as Somewhat closely. This is a measurement issue within to the questions asked in the survey and all self-identifying variables in general. @figure2 shows the quantity of respondents in each group. The most common response is that they follow somewhat closely, and the other responses are relatively even.

@@

Overall, the results of this analysis were inconclusive to measure how income impacts individuals' propensity to follow republican rhetoric. There are many reasons this could be true and as stated in the @intro there are multiple schools of thought previously studied on the topic.

# Discussion {sec-disc}
## Inconclusive – not a futile exercise 
Our results show no evident relationship between income and Republic following. This result is important in of itself – the insignificant relationship suggests income does not influence an individual’s choice of presidential candidate. There can be many possible reasons. The first reason is that other factors apart from income influence this choice to a much greater extent. For example, a family of 5 kids may be more influenced by a candidate promising better institutional infrastructure like more schools and subsidized education. As mentioned in ‘Factors Influencing Voting Decision: A Comprehensive Literature Review’ [@mdpi], the main headings influencing decisions include ‘demographic factors, psychological factors, sociocultural factors and economic factors.’ Income is only one of the many economic factors, and so may have a very small influence on voting preference.

@@

# Appendix {-}
## Figures

>>>>

---
title: "How Household Income Impacts Political Views, Focusing on Mr. Donald Trump, Yielded Inconclusive Results"
author: 
  - Gavin Crooks
  - Samarth Rajani
thanks: "Code is available at: LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media. Data is available at :LINK.https://www.aeaweb.org/articles?id=10.1257/aer.20190658&ArticleSearch%5Bwithin%5D%5Barticletitle%5D=1&ArticleSearch%5Bwithin%5D%5Barticleabstract%5D=1&ArticleSearch%5Bwithin%5D%5Bauthorlast%5D=1&ArticleSearch%5Bq%5D=&JelClass%5Bvalue%5D=I3&journal=1&page=2&per-page=21&from=j"
date: today
date-format: long
abstract: "This study explores the relationship between household income, ethnicity, and political preferences using data from 'The Welfare Effects of Social Media.' Despite inconclusive findings challenging the presumed impact of income on political views, our innovative metric, measuring respondents' propensity to follow Donald Trump, did not provide conclusive results. This introduces new avenues in political science and statistics, encouraging further exploration into understanding the dynamics of these relationships. The study contributes to the ongoing discourse on the complex relationships between income, ethnicity, and political inclinations." 
format: pdf
number-sections: true
toc: true
bibliography: references.bib

@@

Our data has been obtained from ‘The Welfare Effects of Social Media’ [@paper]. To obtain the raw data please use thine links available in the repo or bibliography. 
Our code is supported by the following packages [@citeR] [@tidy] [@her] [@knitr]

@@

The data for this paper was collected from the replication package of the paper ‘The Welfare Effects of Social Media’[@paper] The authors from that data had collected this data themselves using an online survey platform called Qulatrics, inquiring about personal information such as personal names, IP addresses, extent of by which the subjects follow politics, etc. Notably the dataset would contain a lot of confidential information, that if released in the replication package would cause ethical problems. As a result, the authors included the de-identified versions of their data collected, which was used in our analysis too.This method of data collection obtained a large amount of data, 24 000 responses, but will bring to question the quality of the data. All responses were self reported, imputing a form of bias as the values obtained my not be the true. Specifically terms used such as 'Somewhat closely' is up into interpretation. THis is explored with in each variable and tested using the figures below.

@@

Another variable of use is the extent to which the subjects follow Mr. Donald Trump, leader of the Republican Party. The possible responses were ‘Not at all closely', 'Somewhat closely,’ ‘Rather closely’ and ‘Very Closely,’ and the respondents could select one of these options which will become our measure of measuring subscription to Republican ideas. This data is represented using a pie chart in @fig-figure2. This variable faces many problems as this categorical scale is not consistent across respondents. To be specific we mean someone who identifies as someone who does follows Not at all closely can be following Mr. Donald Trump more than someone who identifies as Somewhat closely. This is a measurement issue within to the questions asked in the survey and all self-identifying variables in general. @fig-figure2 shows the quantity of respondents in each group. The most common response is that they follow somewhat closely, and the other responses are relatively even.

@@

Overall, the results of this analysis were inconclusive to measure how income impacts individuals' propensity to follow republican rhetoric. There are many reasons this could be true and as stated in the @sec-intro there are multiple schools of thought previously studied on the topic.

# Discussion {#sec-disc}
## Inconclusive – not a futile exercise 
Our results show no evident relationship between income and Republic following. This result is important in of itself – the insignificant relationship suggests income does not influence an individual’s choice of presidential candidate. There can be many possible reasons. The first reason is that other factors apart from income influence this choice to a much greater extent. For example, a family of 5 kids may be more influenced by a candidate promising better institutional infrastructure like more schools and subsidized education. As mentioned in ‘Factors Influencing Voting Decision: A Comprehensive Literature Review’ [@mdpi], the main headings influencing decisions include ‘demographic factors, psychological factors, sociocultural factors and economic factors.’ Income is only one of the many economic factors, and so may have a very small influence on voting preference.

@@

# Appendix {#sec-appendix}
